humanCdGF_predict_cfu

### 2016-10-11 code/HumanCdGF_predict_cfu.Rmd
add plot to compare median colonization to day 0

---
title: "humanCdGF_predict_cfu"
author: "Nicholas Lesniak"
date: "October 7, 2016"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE, fig.width=10, fig.height=5)
```

```{r load packages}
pack_used <- c('randomForest','miscTools','tidyr','ggplot2','dplyr', 'pROC', 'knitr', 'cowplot')
for (dep in pack_used){
  if (dep %in% installed.packages()[,"Package"] == FALSE){
    install.packages(as.character(dep), repos = 'http://cran.us.r-project.org', 
                     quiet=TRUE);
  }
  library(dep, verbose=FALSE, character.only=TRUE)
}
```

@@

seed <- 1
n_trees <- 2001
rf_plot_list <- list()
n_features <- 10
cutoff <- 30

@@

>>>>

---
title: "humanCdGF_predict_cfu"
author: "Nicholas Lesniak"
date: "October 7, 2016"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE, fig.width=7, fig.height=8)
```

@@

seed <- 1
n_trees <- 1001
rf_plot_list <- list()
n_features <- 10
cutoff <- 30

@@

### Reframing the question for predicting CFU
#### 1. Can Day 0 community predict the median steady-state-colonized CFU?
#### 2. Can Day 0 community predict the rate of colonization?

Cage_dep_df <- meta_data %>% 
  filter(day == 0) %>% 
  select(human_source, sample_id, cage_id) %>% 
  left_join(rel_abund_shared_01)
Predict_cage_df <- select(Cage_dep_df, -human_source, -sample_id)
Predict_source_df <- select(Cage_dep_df, -cage_id, -sample_id)
Predict_cage_df$cage_id <- factor(Predict_cage_df$cage_id)
Predict_source_df$human_source <- factor(Predict_source_df$human_source)

# identify OTUs that predict cage/inocula
set.seed(seed)
rf_cage <- randomForest(cage_id ~ ., data=Predict_cage_df, importance=TRUE, ntree=n_trees)
set.seed(seed)
rf_source <- randomForest(human_source ~ ., data=Predict_source_df, importance=TRUE, ntree=n_trees)
features_source <- head(sort(rf_source$importance[ , 'MeanDecreaseAccuracy'], decreasing = T), 65)
features_cage <- head(sort(rf_cage$importance[ , 'MeanDecreaseAccuracy'], decreasing = T), 65)
features_all <- unique(c(names(features_cage), names(features_source)))

colonized_cfu <- meta_data %>% 
  filter(day %in% c(4:10), !cage_id == 'NP1', 
         Early_Euth == FALSE, cdiff_strain == 431) %>% 
  select(mouse_id, cage_id, log_cfu) %>% 
  group_by(mouse_id, cage_id) %>% 
  summarise(log_cfu_med = median(log_cfu)) %>% 
  ungroup()
sample_list_d0 <- shared_list_days %>% 
    filter(day == 0) %>% 
    select(-day)
community_d0.df <- rel_abund_shared_01 %>% 
    right_join(sample_list_d0) %>% 
    select(-sample_id)
cfu_shared_colonized.df <- colonized_cfu %>% 
    right_join(community_d0.df) %>% 
    select(-mouse_id, -cage_id, -one_of(features_all)) %>% 
    filter(!is.na(log_cfu_med))

test <- Boruta(log_cfu_med ~ ., data = cfu_shared_colonized.df, maxRuns = 5000)

cfu_shared_colonized <- cfu_shared_colonized.df %>% 
  select(log_cfu_med, one_of(names(test$finalDecision[test$finalDecision %in% c('Confirmed', 'Tentative')])))

set.seed(seed)
rf_all <- randomForest(log_cfu_med ~ ., data = cfu_shared_colonized.df,
                           importance = TRUE, ntree = n_trees, mtry = 14)
set.seed(seed)
rf_bor <- randomForest(log_cfu_med ~ ., data = cfu_shared_colonized,
                           importance = TRUE, ntree = n_trees, mtry = 2)

plot(importance_sorted/max(importance_sorted))
importance_sorted <- sort(importance(rf_bor)[,1], decreasing = T)
top_important_OTU <- rownames(data.frame(importance_sorted[importance_sorted/max(importance_sorted) > 0.6]))

set.seed(seed)
rf_n_features <- randomForest(log_cfu_med ~ ., data=cfu_shared_colonized.df %>% 
                                    select(log_cfu_med, one_of(top_important_OTU)),
                                           importance=TRUE, ntree=n_trees, mtry = 2)

importance_final <- sort(importance(rf_n_features)[,1], decreasing = T)

rf_prediction <- data.frame(observed = cfu_shared_colonized.df$log_cfu_med, 
                       predicted = rf_n_features$predicted)
rf_prediction <- merge(rf_prediction, colonized_cfu, by.x = 'observed', by.y = 'log_cfu_med')
predict_plot_day <- 
  ggplot(data = rf_prediction, aes(x=observed, y=predicted, color = cage_id)) + geom_point() + 
    labs(x= 'Observed CFU (log10)', y = 'Predicted CFU (log10)',
         title = "Random Forest Prediction of Colonized C Difficile CFU from Day 0 Community") + 
    theme_bw() + xlim(0,9) + ylim(0,9) + 
    annotate('text', x=0.15*(max(rf_prediction$observed)), 
             y=0.95*(max(rf_prediction$predicted)), 
             label=paste0('R^2 = ', round(tail(rf_n_features$rsq, 1), 2)))

importance.df <- data.frame(otu = names(importance_final),
                            importance = importance_final)
otu_taxa <- get_tax(1, importance.df$otu, tax_file)

importance_plot_day <- 
  ggplot(data = importance.df, aes(x = factor(otu), y = importance)) + 
        geom_point() + scale_x_discrete(limits = rev(importance.df$otu),
                                        labels = rev(paste(otu_taxa[,1],' (',
                                                       rownames(otu_taxa),')',
                                                       sep=''))) +
        labs(x= '', y = '% Increase in MSE') + theme_bw() + coord_flip()
